---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(gcloud)
new_server <- "r-snyhepwjps"
```
# gcloud

This package provides a light wrapper to Google's *gcloud* CLI.  It currently focuses on automating the creation and setup of VM instances.

## Installation

The Google SDK needs to be installed prior use of this package: https://cloud.google.com/sdk/install

``` r
devtools::install_github("edgararuiz/gcloud")
```

## Basics

Use `gcloud_shell()` to run standard commands against your current Google Cloud project.  The function automatically prefixes *gcloud* to the command. 

```{r}
library(gcloud)

gcloud_shell("compute instances list")
```

`gcloud_to_tibble()` parses some of the CLI's output into tibbles. It automatically detects the location and length of each column.

```{r}
gcloud_to_tibble("compute instances list")
```

Higher level wrappers are available as well, such as `gcloud_instances()`, `gcloud_images()` and `gcloud_machine_types()`.

```{r}
gcloud_instances()
```

To see the current configuration settings for your user use `gcloud_config()`.

```{r}
gcloud_config()
```

## Server creation and configuration

A new server VM instance can be created using `gcloud_new_instance()`.  The function contains enough default arguments that allows the creation of a 8 CPU server with Ubuntu 18.04 by just calling the function.

```{r, eval = FALSE}
new_server <- gcloud_new_instance()
```


Once created, it is easy to interact with the new server using 

```{r}
gcloud_run("ls /home", server_name = new_server)
```

Another way to interact and configure the server is to copy and run a bash file. This package comes with an example shell file.

```{r}
script <- system.file("shell/revdep.sh", package = "gcloud")

head(readLines(script, encoding = "UTF-8"), 12)
```

The `revdep.sh` shell file will do the following:

- Installs Linux dependencies, including Java
- Install the preview version of RStudio Server Open Source
- Builds R from source
- Creates a user named `rstudio` and sets the password to `rstudio`
- Pre installs `devtools` and `revdepcheck`

Use `gcloud_sh_run()` to run the shell file.  It will copy and execute the script for you.

```{r, eval = FALSE}
gcloud_sh_run(script, server_name = new_server)
```

## Full example

```{r, eval = FALSE}
repo <- "edgararuiz/modeldb"
# Create a new instance
new_server <- gcloud_new_instance()
Sys.sleep(30)
# Copy and run the shell script
gcloud_sh_run(
  system.file("shell/revdep.sh", package = "gcloud"), 
  server_name = new_server)
# Clone the repo, under the rstudio user
gcloud_run(
  paste0("git clone https://github.com/", repo,".git"), 
  new_server, 
  "rstudio")
# Open the RStudio port access
gcloud_shell(
  "compute firewall-rules create rstudio --rules tcp:8787 --action allow"
  )
# Get the instance list in order to get the external IP
instances <- gcloud_instances()
ext_ip <- instances$EXTERNAL_IP[instances$NAME == new_server]
# Automatically navigate to the RStudio UI
browseURL(paste0("http://", ext_ip, ":8787")) 
```

